Plant-pollintor networks worldwide are composed by the same specific building

### 2021-09-28 Draft/Draft.Rmd
Caption figure 2

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks (Lanuza et al., unpublished). Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) over and under-represented motifs; (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

@@

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonmoic rank level and not with functional traits because the main orders of floral visitors differed clearly in form and behaviour and had lower lower complexity of higher taxnomic ranks than plants (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera, (vi) Coleoptera, (vii). Further, other 3 functional groups (i.e., lizards, birds and other insects) were considered for analysis but excluded from results because of their low representation in the set of networks.

@@

The hierarchical cluster analysis divided the dataset into five different clusters with different and overlapping characteristics (**Figure S3** and **Figure S4**). Cluster 1 referred to as 'selfing herbs' consisted on herbs with hermaphroditic flowers with high levels of autonomous selfing. Cluster 2 or 'small outcrossing perennials' had small perennial species with a mixed of life forms (i.e., trees shrubs and herbs) with outcrossing hermaphroditic flowers. Cluster 3 referred to as 'self-incompatible perennials with large flowers' comprised perennial species with a mixed of life forms and large self-incompatible hermaphroditic flowers with high number of ovules. Cluster 4 or 'tall plants with small unisexual flowers' had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding systems with small flowers but the highest numbers of flowers and inflorescences per plant. Finally, cluster 5 or 'short-lived outcrossers with long zygomorphic flowers' consisted on small perennial and short-lived herbs with long self-compatible zygomorphic flowers that were unable to self-pollinate.

In total, there were 1126 species of floral visitor with 6325 interactions recorded with plants. We divided them into 6 main different taxonomical groups: bees with 2256 interactions with plants, non-syrphid Diptera with 1768 interactions, syrphids with 845 interactions, Lepidoptera with 437 interactions, Coleoptera with 432 interactions and non-bee Hymenoptera with 362 interactions.

@@

Most motifs were under- and over-represented (close to the 1st and 99th percentile, respectively) in the comparison between empirical and simulated networks (**Figure 2**). Further, the exclusion of singletones did not change these results (**Figure S1**). Motifs 3, 5, 9, 10 and 14 were under-represented in empirical networks, that is, all under the 25th percentile and the closest motifs to the 1st percentile. Remarkably, four out of five of these motifs belonged to the largest path length classification (i.e., core-peripheral). In addition, motifs 2, 6, 7, 16 and 17 were over-represented, all over the 75th percentile and thus the closest to the 99th percentile. In contrast to the under-represented motifs, over-represented motifs belonged to the two shortest path length groups (i.e., complete and fan). The remaining motifs (i.e., 4, 8, 11, 12, 13, and 15) were between the 25th and 75th percentile but closest to over-representation.

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=9,fig.height=5, fig.cap="**Figure 2.** Comparison of the motif frequencies between empirical and simulated networks. Average percentages of motifs close to 0 and 100 indicate under- and over- representation in empirical networks. The different motifs are coloured by the mean motif path length as done in Simmons et al. (2020)."}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
d <- add_study_id(d)
#str(d)
#head(d)

@@

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(y=as.factor(motif), x=100*percentil_sizeclass, 
                 color = as.factor(Broad_categories)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = motif_means,aes(y = as.factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(y = as.factor(motif), x=100*mean), 
                size=2)+
  labs(y="Two to five node motifs", x = "Percentile",colour="Path length classification")+
  theme_bw() + scale_color_manual(labels = c('Strong' = expression(paste("Complete (", bar(x), " = 1.38)")),'Fan' = expression(paste("Fan (", bar(x), " = 1.48)")), 'Medium-weak' = expression(paste("Asymmetric complete (", bar(x), " = 1.60)")), "Weak" =expression(paste("Core-peripheral (", bar(x), " = 1.85)"))), name = "Path length classification",values = c("grey25","springgreen4","sienna2","goldenrod2"),guide = guide_legend(title.position = "top",ncol = 1)) +
theme(legend.text.align = 0)+
  theme(legend.position="right")

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=8,fig.width=8,fig.align = 'center', fig.cap="\\textbf{Figure 3.} Heatmap indicating under- and over- representation of pollinator and plant functional groups in the different motif positions. The different motif positions are dividied by the average path length clasification by @simmons2020."}

@@

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(y=as.factor(motif), x=100*percentil_sizeclass, 
                 color = as.factor(Broad_categories)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = motif_means,aes(y = as.factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(y = as.factor(motif), x=100*mean), 
                size=2)+
  labs(y="Motif", x = "Percentile (%)",colour="Path length classification")+
  theme_bw() + scale_color_manual(labels = c('Strong' = expression(paste("Complete (", bar(x), " = 1.38)")),'Fan' = expression(paste("Fan (", bar(x), " = 1.48)")), 'Medium-weak' = expression(paste("Asymmetric complete (", bar(x), " = 1.60)")), "Weak" =expression(paste("Core-peripheral (", bar(x), " = 1.85)"))), name = "Path length classification",values = c("grey25","springgreen4","sienna2","goldenrod2"),guide = guide_legend(title.position = "top",nrow = 2)) +
theme(legend.text.align = 0)+
  theme(legend.position="bottom")

>>>>

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks. Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) over and under-represented motifs; (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

@@

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera, (vi) Coleoptera, (vii). However, a minor set of species belonged to other functional groups that were considered in the analysis as lizards, birds and other insects but not discussed further because of their low representation in the set of analysed networks.

@@

The hierarchical cluster analysis divided the dataset into five different clusters with different and overlapping characteristics (**Figure S3** and **Figure S4**). Cluster 1 referred to as 'selfing herbs' consisted on herbs with hermaphroditic flowers with high levels of autonomous selfing. Cluster 2 or 'small outcrossing perennials' had small perennial species with a mixed of life forms (i.e., trees shrubs and herbs) with outcrossing hermaphroditic flowers. Cluster 3 referred to as 'self-incompatible perennials with large flowers' comprised perennial species with a mixed of life forms and large self-incompatible hermaphroditic flowers with high number of ovules. Cluster 4 or 'tall plants with small unisexual flowers' had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding systems, small flowers and the highest numbers of flowers and inflorescences per plant. Finally, cluster 5 or 'short-lived outcrossers with long zygomorphic flowers' consisted on small perennial and short-lived herbs with long self-compatible zygomorphic flowers that were unable to self-pollinate.

In total, there were 1126 species of floral visitor with 6325 interactions recorded with plants. From highest to lowest records of interactions of floral visitors with plants we found: bees with 2256 interactions, non-syrphid Diptera with 1768 interactions, syrphids with 845 interactions, Lepidoptera with 437 interactions, Coleoptera with 432 interactions and non-bee Hymenoptera with 362 interactions.

@@

Most motifs were under- and over-represented (close to the 1st and 99th percentile, respectively) in the comparison between empirical and simulated networks (**Figure 2**). Motifs 3, 5, 9, 10 and 14 were under-represented in empirical networks, that is, all were close to the 1st percentile and under the 25th percentile. Remarkably, four out of five of these motifs belonged to the largest path length classification (i.e., core-peripheral). In addition, motifs 2, 6, 7, 16 and 17 were over-represented, all over the 75th percentile and thus the closest to the 99th percentile. In contrast to the under-represented motifs, over-represented motifs belonged to the two shortest path length groups (i.e., complete and fan). The remaining motifs (i.e., 4, 8, 11, 12, 13, and 15) were between the 25th and 75th percentile but closest to over-representation. Remarkably, the exclusion of singletones from this analysis did not change the observed results (**Figure S1**).

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=9,fig.height=5, fig.cap="**Figure 2.** Comparison of motif frequencies between empirical and simulated networks. This is shown with the mean percentage of motif frequencies in empirical networks that were over the motif frequencies of the simulated ones (percentiles). This was done by network (coloured dots) and then averaged for all networks (black dots with error bars that correspond to the standard deviation). The different motifs are coloured by the four categories of path length determined in Simmons et al. (2020)."}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
d <- add_study_id(d)
#str(d)
#head(d)

@@

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(y=as.factor(motif), x=100*percentil_sizeclass, 
                 color = as.factor(Broad_categories)),
             position = "jitter",alpha=0.8)+
  geom_errorbar(data = motif_means,aes(y = as.factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(y = as.factor(motif), x=100*mean), 
                size=2)+
  labs(y="Two to five node motifs", x = "Percentile",colour="Path length classification")+
  theme_bw() + scale_color_manual(labels = c('Strong' = expression(paste("Complete (", bar(x), " = 1.38)")),'Fan' = expression(paste("Fan (", bar(x), " = 1.48)")), 'Medium-weak' = expression(paste("Asymmetric complete (", bar(x), " = 1.60)")), "Weak" =expression(paste("Core-peripheral (", bar(x), " = 1.85)"))), name = "Path length classification",values = c("grey36","lightblue3","navajowhite2","darkorange4"),guide = guide_legend(title.position = "top",ncol = 1)) +
theme(legend.text.align = 0)+
  theme(legend.position="right")
#Trying to maintain a gradient from low to high in colors

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=8,fig.width=8,fig.align = 'center', fig.cap="\\textbf{Figure 3.} Heatmap indicating under- and over- representation of pollinator and plant functional groups in the different motif positions. The different motif positions are dividied by the average path length clasification determined by Simmons et al. (2020)."}

@@

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(y=as.factor(motif), x=100*percentil_sizeclass, 
                 color = as.factor(Broad_categories)),
             position = "jitter",alpha=0.8)+
  geom_errorbar(data = motif_means,aes(y = as.factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(y = as.factor(motif), x=100*mean), 
                size=2)+
  labs(y="Two to five node motifs", x = "Percentile",colour="Path length classification")+
  theme_bw() + scale_color_manual(labels = c('Strong' = expression(paste("Complete (", bar(x), " = 1.38)")),'Fan' = expression(paste("Fan (", bar(x), " = 1.48)")), 'Medium-weak' = expression(paste("Asymmetric complete (", bar(x), " = 1.60)")), "Weak" =expression(paste("Core-peripheral (", bar(x), " = 1.85)"))), name = "Path length classification",values = c("grey36","lightblue3","navajowhite2","darkorange4"),guide = guide_legend(title.position = "top",ncol = 1)) +
theme(legend.text.align = 0)+
  theme(legend.position="right")
